---
title: "Electricity Usage in Toronto’s Public Facilities: Analyzing the Impact of Building Size on Energy Consumption"
author: 
  - Irene Liu
thanks: "Code and data are available at: https://github.com/zilin1017/Toronto_annual_energy_consumption.git."
date: today
date-format: long
abstract: "This study examines the relationship between building size(m²) and electricity consumption(kWh) in Toronto's public facilities. Data from over 500 buildings, including schools,libraries, and hospitals, show a consistent positive correlation between GFA and eletricity usage. Larger facilities, particularly healthcare centers, demonstrate higher energy consumption per square meter. These results help inform efforts to improve energy management in Toronto’s public buildings."
format: pdf
toc: true
number-sections: true
bibliography: references.bib
---

```{r}
#| include: false
#| warning: false
#| message: false

##install the packages
# install.packages("opendatatoronto")
# install.packages("tidyverse")
# install.packages("knitr")

library(tidyverse)
library(knitr)
library(dplyr)
library(ggplot2)
library(readr)


```


# Introduction

Public infrastructure, including schools, libraries, healthcare centers, and government offices, forms the backbone of urban services in any modern city. As Toronto continues to expand, managing electricity usage in public buildings becomes increasingly critical. Energy efficiency in public facilities not only reduces operational costs but also contributes to environmental goals, such as carbon reduction and climate change mitigation[@droege2018urban].

One of the primary determinants of electricity consumption is building size, commonly measured as Gross Floor Area (GFA). Larger buildings tend to consume more energy for lighting, heating, cooling, and equipment operations, but this relationship is influenced by other factors such as building type, operational hours, and energy management systems [@hong2016occupant]. As such, it is crucial for policymakers to understand how building size and other factors influence electricity consumption in Toronto’s public facilities [@li2009managing].

Several studies have shown that energy efficiency measures in public buildings can lead to significant reductions in electricity consumption. For example, retrofitting large government buildings or healthcare facilities with modern energy management systems can reduce energy use, while smaller buildings like libraries and schools often exhibit lower per-square-meter energy consumption [@scott2008impact].

This paper explores the relationship between Property GFA (m²) and electricity usage (kWh) in over 500 public facilities in Toronto. The study aims to identify patterns in electricity consumption and provide insights for optimizing energy use in Toronto's public sector.


# Data {#sec-data}

## Raw data
The data used in this paper came from the OpenData Toronto portal through the library [@opendt]. Data were cleaned and analyzed using the open source statistical programming language R [@citeR]. The particular data set used to analyze the Consumption of Energy is the Annual Energy Consumption Data in 2021 [@2021datatoronto]. The following packages helped with the data analysis: tidyverse [@tidyverse], here [@here]), dplyr [@dplyr] and ggplot2 [@ggplot] was use for making graphics. All data analysis was conducted using R [@citeR].

The data used in this analysis comes from the City of Toronto's Open Data Portal[@opendt], specifically focusing on energy consumption across various property types in Toronto for the year 2021[@2021datatoronto]. The dataset consists of 1,802 unique entries, each representing a specific property in various regions of Toronto. For each property, the data captures critical metrics such as the building size, self-reported in gross floor area (GFA) in square meters (m²), and the total electricity consumption, measured in kilowatt-hours (kWh), purchased from the grid. No personal or property-specific information is included. The data was last updated on July 4, 2023, and helps understand energy usage trends across different building types in Toronto.

## Cleaned Data

The cleaning process focused on removing irrelevant or incomplete data to ensure the dataset was suitable for analyzing electricity consumption patterns in Toronto's public sector. Missing values(NA) were removed during the data cleaning process, particularly those related to electricity usage. And buildings that did not meet the size criteria (less than 30 square meters or greater than 2,500 square meters) were also filtered out.Additionally, variables irrelevant to the study’s focus were also omitted. Furthermore, the cleaned data retains only the key variables required for the analysis: property type, electricity use, floor area, and log-transformed area. This ensures that the final dataset is both accurate and relevant for understanding electricity consumption trends across different public facilities in Toronto.

A sample of the cleaned data can be seen in @tbl-1. The cleaned analysis dataset is loaded using the R programming language [@citeR]. The folder structure for this analysis is also inherited from Professor Rohan Alexander [@rohan]. Figures and tables are then generated using the knitr [@knitr] and tidyverse [@tidyverse] packages.

```{r}
#| label: tbl-1
#| tbl-cap: Sample of cleaned data
#| echo: false

analysis_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data_subset <- analysis_data[ c("Property_Type", "kWh", "Area", "log_Area")]

filtered_data <- analysis_data_subset|>
  filter(!is.na(kWh) & kWh != "NA") |>   
  distinct(Property_Type, .keep_all = TRUE) 

kable(
  head(filtered_data, 6), 
  col.names = c("Property Type", "Electricity Use (kWh)", "Area (m²)", "Log of Area (m²)"),
  align = c("l", "c", "c", "c","c")
)

```

## Data Analysis
In  @fig-Area, we observe the density of electricity usage versus building size across different cities in Toronto. Denser clusters of data points in central areas such as Toronto and Scarborough suggest a higher concentration of both building sizes and electricity consumption. Specifically, Toronto shows a dense clustering of buildings in the range of 25,000 to 50,000 m², indicating that larger facilities in central locations tend to consume significantly more electricity compared to those in peripheral cities. In contrast, cities like Maple and York have relatively fewer buildings, and their electricity consumption is more evenly distributed, as indicated by the sparse density of points. This density pattern supports the paper's objective of analyzing how building size influences electricity use and points to central urban areas as having the greatest potential for energy efficiency interventions.

```{r}
#| label: fig-Area
#| fig-cap: Density Plot for Property Area in Toronto (By City)
#| echo: false
#| message: false
#| warning: false


raw_data <- read_csv(here::here("data/raw_data/raw_data.csv"), skip=5)
analysis_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))

annual_energy_consumption_data_2021 <- raw_data |>
  dplyr::select(`Property GFA - Self-Reported (m²)`, `Electricity Use - Grid Purchase (kWh)`,  `City`)|>
  dplyr::mutate(Area = `Property GFA - Self-Reported (m²)`,
                kWh = `Electricity Use - Grid Purchase (kWh)`,
                City = `City`) 


annual_energy_consumption_data_2021 %>%
  ggplot(aes(x = Area, y = kWh, color = City)) +
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE) +
  facet_wrap(~ City, scales = "free") + 
  ggtitle("Electricity Usage vs Building Size by City") +
  xlab("Building Size (GFA in m²)") +
  ylab("Electricity Usage (kWh)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5,  size = 10),
    axis.ticks.y = element_blank(),  
    axis.text.y = element_blank(),
    axis.text.x = element_text(angle = 45, hjust = 1, size=4.5),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
  )

```

```{r}
#| label: fig-Property
#| fig-cap: Percentage Electricity Usage by Properties 
#| echo: false
#| message: false
#| warning: false

raw_data <- read_csv(here::here("data/raw_data/raw_data.csv"), skip=5)
analysis_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))

analysis_data <- analysis_data |>
  mutate(Property_Type = case_when(
    grepl("Other", `Property_Type`) ~ "Other",
    TRUE ~ `Property_Type`
  ))
kWh_count <- analysis_data |>
  group_by(Property_Type)|>
  summarise(kWh = n()) |>
  arrange(desc(kWh))

ggplot(kWh_count, aes(x = Property_Type, y = kWh, fill = Property_Type)) +
  geom_bar(stat = "identity") +
  geom_text(aes(label = paste0(round(kWh/sum(kWh) *100 , 1), "%")), vjust=0.2, size = 2) +
  ggtitle("Total Electricity Usage by Property Type") +
  xlab("Property Type") +
  ylab("Total Electricity Usage (kWh)") +
  theme_minimal() +
  theme(
    plot.title = element_text(hjust = 0.5, margin = margin(b = 45), size = 8),
    axis.text.y = element_blank(), 
    axis.ticks.y = element_blank(), 
    axis.text.x = element_text(angle = 65, hjust = 1, size=6),
    legend.key.size = unit(0.3, "cm"),
    legend.text = element_text(size = 5),
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
    )
```

```{r}
#| label: fig-log-area
#| fig-cap: Density Plot of Log_Area
#| echo: false
#| message: false
#| warning: false


analysis_data |>
  ggplot(aes(x=log_Area)) +
  geom_histogram(aes(y= ..density.., fill = "pink")) +
  geom_density() +
  ggtitle("Density Plot for Property Area in Toronto") +
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
    )
  
```

```{r}
#| label: fig-kWh
#| fig-cap: Density Plot of kWh
#| echo: false
#| message: false
#| warning: false


analysis_data <- analysis_data |> 
  dplyr::filter(kWh > 0)

analysis_data |>
  ggplot(aes(x = kWh)) +
  geom_histogram(aes(y = ..density..), fill = "blue", binwidth = 250000, alpha = 0.6) +  
  geom_density() +
  ggtitle("Density Plot for Electricity Use in Toronto") +
  theme_minimal()+
  theme(
    panel.grid.major = element_blank(), 
    panel.grid.minor = element_blank()
    )

```
In @fig-Property, the bar plot illustrates the total electricity usage across various property types. Notably, the Other category dominates with 38.3% of the total electricity consumption, suggesting a wide range of public facilities categorized under this label that may significantly differ in size and function. Other high-consumption categories include Fire Stations (15.8%), Libraries (13.7%), and Social/Meeting Halls (11.3%). These property types account for a substantial portion of electricity use, highlighting their operational demands.
The distribution of percentages provides a clear distinction between high-demand and low-demand property types. This segmentation supports the paper's objective of identifying the specific categories where energy efficiency efforts may yield the most impact, particularly in facilities that operate intensively throughout the year.

Density Plot for Property Area in Toronto @fig-log-area: it illustrates the distribution of property sizes in logarithmic form, showing that most buildings fall within a moderate size range. However, a smaller number of properties have much larger areas, indicated by the tail on the right side of the plot. This observation illustrates that the larger buildings tend to consume more electricity, but the relationship is not strictly proportional across the full range of building sizes.

The density plot for electricity @fig-kWh use shows a sharp peak in the lower range of electricity consumption, with most facilities consuming less than 5,000,000 kWh annually. This suggests that the majority of public facilities operate with relatively low energy demands. However, the right tail extends toward much higher consumption values, with a few facilities reaching close to 20,000,000 kWh. The skewness of this distribution indicates the presence of outliers—facilities that consume significantly more energy than others. These outliers may represent large, energy-intensive buildings or those with specific operational requirements.
\newpage

# Results
```{r}
#| label: fig-modelresults
#| fig-cap: Regression of Electricity Use on Area
#| echo: false
#| message: false
#| warning: false


raw_data <- read.csv(here::here("data/raw_data/raw_data.csv"), skip=5)
analysis_data <- read.csv(here::here("data/analysis_data/analysis_data.csv"))


annual_energy_consumption_data_2021 <- analysis_data |>
  mutate(log_Area = log(Area), log_kWh = log(kWh))

ggplot(annual_energy_consumption_data_2021, aes(x = log_Area, y = log_kWh)) +
  geom_point(alpha = 0.5) +
  geom_smooth(method = "lm", se = FALSE, formula = y ~ x, color = "blue") +
  ggtitle("Log-Log Regression of Electricity Use on Area") +
  xlab("Log of Area (m²)") +
  ylab("Log of Electricity Use (kWh)") +
  theme_minimal()

```
Our results, summarized in Figure @fig-modelresults, indicate a positive relationship between electricity use (kWh) and building area (m²) in Toronto's public facilities through log-log regression analysis. This finding aligns with previous research that suggests building size is a key determinant of energy consumption in institutional settings. Larger buildings generally consume more electricity, which is consistent with the basic principle of scaling energy demand with floor area.

However, our analysis suggests that building size alone does not fully explain electricity consumption, indicating that factors such as building age, equipment efficiency, and usage patterns likely play significant roles [@Yohanis2008]. For instance, lighting and heating systems have been shown to contribute disproportionately to energy consumption in larger buildings [@PerezLombard2008]. Consequently, strategies to optimize energy usage should focus not just on large facilities but also on mid-sized properties, which may provide better opportunities for improving efficiency without encountering the high consumption rates associated with the largest properties [@Ouf2017].

By targeting mid-sized buildings for energy efficiency improvements, policy-makers can maximize the return on investment in energy-saving measures, as these properties may experience greater relative reductions in consumption [@Santamouris2017]. Future research should explore the role of operational practices and energy management technologies in reducing consumption across different property sizes.


# Discussion
## First discussion point: Urban Density and Its Impact on Electricity Consumption

@fig-Area highlights how different regions of Toronto exhibit varied patterns of electricity consumption based on their urban density. Cities like Toronto, North York, and Scarborough show denser clusters of data points compared to less populated areas like Maple and Richmond Hill, where fewer observations are present. This difference in data density could suggest that larger, more densely populated urban areas have a higher concentration of public facilities that consume substantial amounts of electricity, supporting previous research [@Ouf2017]. Additionally, the more scattered and sparse data points in regions such as Pickering and Etobicoke indicate that these cities have fewer facilities or public buildings consuming significant electricity, further emphasizing the urban-rural divide in energy demand.

The lack of data points in cities like Maple suggests that the relatively lower number of public buildings could lead to decreased overall electricity usage, which showing that electricity consumption is typically lower in less urbanized areas due to fewer public infrastructures. The data also raises questions about whether the facilities in these regions are less energy-efficient or simply fewer in number, as the lower density could mask potential inefficiencies.

Also, the location of a facility, including its microclimate, can have a substantial impact on energy usage, particularly in cities with diverse weather conditions like Toronto. Buildings in areas with more extreme temperatures (either cold winters or hot summers) require more heating or cooling, which drives up electricity use. This factor is especially relevant for buildings with large window-to-wall ratios or poor insulation, as they tend to lose heat more easily. In addition, the density plots from @fig-Area show that some locations, such as Etobicoke and East York, feature buildings with relatively lower consumption compared to Scarborough or Toronto, indicating that location and local climate could be influencing these trends.

## Second discussion point:Factors Influencing Building Energy Consumption: Property, Functionality, Operational Intensity and Occupancy

Different types of buildings have varying operational needs, which directly impact their energy consumption. For instance, facilities like fire stations and wastewater treatment plants operate 24/7, often requiring extensive heating, ventilation, and cooling systems. In contrast, libraries or office spaces may have more intermittent usage, leading to lower electricity demand. The diversity of energy requirements across property types is a key factor influencing consumption, as demonstrated in @fig-Property where categories such as "Other" and "Fire Stations" stand out with significantly higher usage. 

Buildings with higher occupancy levels or more intense operational schedules tend to consume more energy. Facilities that are open for longer hours or serve a large number of occupants, such as community centers or swimming pools, require more consistent energy to maintain lighting, cooling systems, and equipment. Studies, such as those by [@Yohanis2008], show that buildings with greater occupancy or more energy-intensive activities exhibit higher energy use, even when the building size is relatively modest. In our dataset, "Heated Swimming Pools" and "Social/Meeting Halls" stand out as high-energy users due to their intensive operational demands, despite not being the largest facilities.


## Limitations and Next Steps

While our analysis provides valuable insights into electricity consumption across Toronto’s public facilities, there are several limitations that must be addressed. First, the dataset lacks information on occupancy and operational schedules, both of which are key determinants of energy consumption [@Yohanis2008]. Without these variables, it is difficult to draw firm conclusions about why some buildings consume more energy than others, beyond their size and property type. Additionally, as shown in @fig-log-area and @fig-kWh, the skewed distribution of electricity usage—dominated by a few large consumers—poses challenges for generalizing energy-saving strategies across all building types.

To improve the accuracy and usefulness of the analysis, future studies should incorporate temporal data on electricity use, enabling a better understanding of peak demand periods and seasonal variations in energy consumption. Further, the inclusion of occupancy data and the use of more advanced statistical techniques, such as random forest models or decision trees, could help account for the complex interactions between building size, type, and operational characteristics, leading to more actionable policy recommendations [@Santamouris2017].



\newpage

\appendix

# Appendix {-}
## Dataset and Graph Sketches {-}

Drafts of the necessary dataset and the figures produced by this analysis are available in the other/sketches file of the GitHub repository.

## Attribution Statement {-}
Contains information licensed under the Open Government Licence– Toronto [@open_toronto_license]


\newpage


# References




